---
title: "Tiernet"
author: "Gauthier A."
date: "18/11/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

pacman::p_load(
  rio,        # importing data  
  here,       # relative file pathways
  readxl,     # read excel
  writexl,    # write excel
  #skimr,      # Exploring data
  janitor,    # data cleaning and tables
  lubridate,  # working with dates
  epikit,     # age_categories() function
  tidyverse,  # data management and visualization,
  #gtsummary,  # summary statistics and tests,
  #rstatix,    # summary statistics and statistical tests
  scales,     # easily convert proportions to percents  
  labelled,   # Variable and values labelling
  flextable,  # Format tables,
  sqldf,      #sql queries
  RODBC       # Odbc database connection
)


# use here from the here package
here <- here::here
# use clean_names from the janitor package
clean_names <- janitor::clean_names

# load functions used from R scripts
source_path <- here("R", "RScripts.R")
source(source_path,local = knitr::knit_global())

data_folder <- "data"
image_folder <- "images"
output_folder <- "output"

###############################################################################
#################   I. Importing & Exploring Data         #####################
###################   Version of Tier.net : 1.13.4      #######################
###############################################################################

```

```{r, include=FALSE}

con<-RODBC::odbcDriverConnect("Driver={Microsoft Access Driver (*.mdb)};DBQ=C:/Users/Gauthier Abdala/Documents/RWD/TierNet/data/TierDESKinshasa042022.mdb")

vis_raw <- RODBC::sqlFetch(con,"VIS") #Follow Up clinic Visit table
pat_raw <- RODBC::sqlFetch(con,"PAT") #Socio demographic char. & outcome table
lab_raw <- RODBC::sqlFetch(con,"LAB") #Laboratory data table
dem_raw <- RODBC::sqlFetch(con,"DEM") #Demographic information

#pregnancy_raw <- RODBC::sqlFetch(con,"PREGNANCY")  #Pregnancy information
tb_raw        <- RODBC::sqlFetch(con,"TB")         #Tuberculosis Information
vis_tb_raw    <- RODBC::sqlFetch(con,"VIS_TB")     #TB Visit Table
transfers_raw <- RODBC::sqlFetch(con,"TRANSFERS")  #Transfer In/Out Table
#vis_preg_raw  <- RODBC::sqlFetch(con,"VIS_PREG")   #Antenatal Care
#infant_raw    <- RODBC::sqlFetch(con,"INFANT")     #Labour and Delivery


#############################################################################
########################      II. Cleaning data    ##########################
#############################################################################


############# 1. Cleaning column names    #################
###########################################################

vis <- vis_raw %>% janitor::clean_names()
pat <- pat_raw %>% janitor::clean_names()
lab <- lab_raw %>% janitor::clean_names()
dem <- dem_raw %>% janitor::clean_names()

#pregnancy <- pregnancy_raw  %>% janitor::clean_names()
tb        <- tb_raw         %>% janitor::clean_names()
vis_tb    <- vis_tb_raw     %>% janitor::clean_names()
transfers <- transfers_raw  %>% janitor::clean_names()
#vis_preg  <- vis_preg_raw   %>% janitor::clean_names()
#infant    <- infant_raw     %>% janitor::clean_names()


####### 2. Select,add and/or re order/rename columns #######
############################################################
vis <- vis %>%
  select(patient, visit_dmy, tb_status, who_stage, next_visit_dmy)

pat <- pat %>%
  select(patient, cohort, facility, birth_dmy,gender, frsvis_dmy,hivp_dmy, haart,
         haart_dmy, fhv_stage_who, tb_fhv, method_into_art, transfer_in_dmy, outcome,
         outcome_dmy)

lab <- lab %>%
  select(patient, lab_dmy, lab_id, lab_v, lab_t, episode_type, episode_id)

dem <- dem %>%
  select(patient, folder_number)

tb <- tb %>%
  select(patient, reg_dmy, tb_start_dmy, tb_end_dmy, tb_outcome, episode_id)

vis_tb <- vis_tb %>%
  select(patient, visit_dmy, next_visit_dmy, episode_id)

transfers <- transfers %>%
  select(patient, episode_id, facility_from, facility_to, transfer_dmy, transfer_action)



################ 3. Cleaning empty/null values     ##########
#############################################################

vis <-trim_data_columns(vis)
vis[is.null(vis)  | vis == "NULL" | vis == ""] <- NA

pat <-trim_data_columns(pat)
pat[is.null(pat)  | pat == "NULL" | pat == ""] <- NA

lab <-trim_data_columns(lab)
lab[is.null(lab)  | lab == "NULL" | lab == ""] <- NA

dem <-trim_data_columns(dem)
dem[is.null(dem)  | dem == "NULL" | dem == ""] <- NA

tb <-trim_data_columns(tb)
tb[is.null(tb)  | tb == "NULL" | tb == ""] <- NA

vis_tb <-trim_data_columns(vis_tb)
vis_tb[is.null(vis_tb)  | vis_tb == "NULL" | vis_tb == ""] <- NA

transfers <-trim_data_columns(transfers)
transfers[is.null(transfers)  | transfers == "NULL" | transfers == ""] <- NA


################ 4. Convert columns classes      #############
##############################################################
#is.date <- function(x) inherits(x, 'Date') #Used to check of column is date


vis <- vis %>%
  mutate(across(.cols = everything(), .fns = as.character)) %>% #Convert all columns to character class
  mutate(
    visit_dmy = parse_date(visit_dmy,"%Y-%m-%d"),
    next_visit_dmy = parse_date(next_visit_dmy,"%Y-%m-%d"),
    who_stage = parse_number(who_stage)
  ) %>%
  mutate(
    visit_dmy = as.Date(as.numeric(visit_dmy),origin ="1970-01-01"),
    next_visit_dmy = as.Date(as.numeric(next_visit_dmy),origin ="1970-01-01")
    )


pat <- pat %>%
  mutate(across(.cols = everything(), .fns = as.character)) %>% #Convert all columns to character class
  mutate(
    birth_dmy = parse_date(birth_dmy,"%Y-%m-%d"),
    frsvis_dmy = parse_date(frsvis_dmy,"%Y-%m-%d"),
    hivp_dmy = parse_date(hivp_dmy,"%Y-%m-%d"),
    haart_dmy = parse_date(haart_dmy,"%Y-%m-%d"),
    transfer_in_dmy = parse_date(transfer_in_dmy,"%Y-%m-%d"),
    outcome_dmy = parse_date(outcome_dmy,"%Y-%m-%d"),
  ) %>%
  mutate(
    birth_dmy = as.Date(as.numeric(birth_dmy),origin ="1970-01-01"),
    frsvis_dmy = as.Date(as.numeric(frsvis_dmy),origin ="1970-01-01"),
    hivp_dmy = as.Date(as.numeric(hivp_dmy),origin ="1970-01-01"),
    haart_dmy = as.Date(as.numeric(haart_dmy),origin ="1970-01-01"),
    transfer_in_dmy = as.Date(as.numeric(transfer_in_dmy),origin ="1970-01-01"),
    outcome_dmy = as.Date(as.numeric(outcome_dmy),origin ="1970-01-01"),
    )

lab <- lab %>%
  mutate(across(.cols = everything(), .fns = as.character)) %>% #Convert all columns to character class
  mutate(
    lab_dmy = parse_date(lab_dmy,"%Y-%m-%d"),
    lab_v = parse_number(lab_v)
  ) %>%
  mutate(
    lab_dmy = as.Date(as.numeric(lab_dmy),origin ="1970-01-01")
    )


tb <- tb %>%
  mutate(across(.cols = everything(), .fns = as.character)) %>% #Convert all columns to character class
  mutate(
    reg_dmy = parse_date(reg_dmy,"%Y-%m-%d"),
    tb_start_dmy = parse_date(tb_start_dmy,"%Y-%m-%d"),
    tb_end_dmy = parse_date(tb_end_dmy,"%Y-%m-%d"),
  ) %>%
  mutate(
    reg_dmy = as.Date(as.numeric(reg_dmy),origin ="1970-01-01"),
    tb_start_dmy = as.Date(as.numeric(tb_start_dmy),origin ="1970-01-01"),
    tb_end_dmy = as.Date(as.numeric(tb_end_dmy),origin ="1970-01-01"),
    )

vis_tb <- vis_tb %>%
  mutate(across(.cols = everything(), .fns = as.character)) %>% #Convert all columns to character class
  mutate(
    visit_dmy = parse_date(visit_dmy,"%Y-%m-%d"),
    next_visit_dmy = parse_date(next_visit_dmy,"%Y-%m-%d"),
  ) %>%
  mutate(
    visit_dmy = as.Date(as.numeric(visit_dmy),origin ="1970-01-01"),
    next_visit_dmy = as.Date(as.numeric(next_visit_dmy),origin ="1970-01-01")
    )

transfers <- transfers %>%
  mutate(across(.cols = everything(), .fns = as.character)) %>% #Convert all columns to character class
  mutate(
    transfer_dmy = parse_date(transfer_dmy,"%Y-%m-%d")
  ) %>%
  mutate(
    transfer_dmy = as.Date(as.numeric(transfer_dmy),origin ="1970-01-01")
    )

# transfers <- transfers %>%
#   filter(transfer_dmy >as.Date("2017-12-31") & facility_from == "CH KABINDA")


################ 5. Cleaning duplicate rows      #############
##############################################################
pat[duplicated(pat),]
pat<-unique(pat)

vis_duplicate <-vis[duplicated(vis),]
vis<-unique(vis)

lab_duplicate <-lab[duplicated(lab),]
lab<-unique(lab)
#TODO Cleaning of lab data for same exame encoded same date, different values...

tb_duplicate <-tb[duplicated(tb),]
tb<-unique(tb)

vis_tb_duplicate <-vis_tb[duplicated(vis_tb),]
vis_tb<-unique(vis_tb)

transfers_duplicate <-transfers[duplicated(transfers),]
transfers<-unique(transfers)

################# 5. Merge tables                #################
##################################################################
vis_data <- left_join(vis,dem,"patient") %>%
  left_join(pat,"patient")


################# 6. Add variables/values labels    ##############
##################################################################
# labelled::val_labels(vis_data$gender)  <-c("1" = "Male", "2" = "Female", "95" = NA, "99" = NA)
# labelled::val_labels(vis_data$haart)   <-c("0" = "Not on ART", "1" = "on ART")
# labelled::val_labels(vis_data$outcome) <-c("10"= "Deceased", "20" = "In care", "30" = "Transfered Out", "40" = "LTFU", "95" = "Unknown")
# labelled::val_labels(vis_data$method_in_art_named) <-c("0" = "New", "1" = "Transfer In", "88" = NA )



```



```{r include=FALSE}

#############################################################################
##############      III. Analyzing data/presenting    #######################
#############################################################################


#Get visits for patients that date of visit = date of art start date


#Q/ Do we need a time window for initiations or do we keep strict art_start_date==visit_date


###################### CD4 LAB DATA ##############################
##################################################################
#A/ Given that in Tier.net we encode results dates for lab tests, we should have a time window check
#TODO : Need to remove duplicates for CD4 data encoded twice for the same visit


###CD4 Lab Data
lab_data_cd4 <-lab %>%
  filter((lab_id =="CD4A") & !is.na(lab_v)) %>%
  mutate(
    cd4_category = age_categories(
      lab_v,
      breakers = c(0,200,1000)
    )
  )




#vis_cd4 <- left_join(vis_data,lab_data_cd4,c("patient"="patient","visit_dmy"="lab_dmy"))

###!!! Important to clarify CD4 time window (here +-7 days) based on result turnaround time and validity period before requesting
###!!! If there exists two CD4 within the time window, which one to pick? Most recent one?

vis_cd4 <- sqldf(
  "select vis.visit_dmy,vis.patient,vis.next_visit_dmy,vis.folder_number,vis.cohort,vis.facility,vis.birth_dmy,
    vis.gender,vis.frsvis_dmy,vis.haart,vis.haart_dmy,vis.method_into_art,vis.tb_status,vis.transfer_in_dmy,vis.outcome,vis.outcome_dmy,lab.lab_dmy,lab.lab_v,lab.cd4_category from vis_data vis
    left join lab_data_cd4 lab 
      on vis.patient = lab.patient and
        lab.lab_dmy  between (vis.visit_dmy-7)  and (vis.visit_dmy + 7) 
  "
)%>%
  rename(
    cd4_date = lab_dmy,
    cd4_value= lab_v
  )
  
 
#Filter duplicate lab results based on CD4 lab date
vis_cd4<-vis_cd4 %>%
  arrange(desc(cd4_date))

#Unique visits
#vis_cd4_data_dup<-vis_cd4[duplicated(vis_cd4[,1:16]),]
vis_cd4_data<-vis_cd4[!duplicated(vis_cd4[,1:16]),]

#Some duplicate visit data include
#Visits with different lab dates based on time window, need to decide which one to pick (soonest maybe?)
# 
# vis_cd4_data[!duplicated(vis_cd4_data[,1:16]),] %>%
#   arrange(patient)

###################### CRAG LAB DATA ##############################
##################################################################

lab_data_crag <-lab %>%
  filter((lab_id =="CRAG") & !is.na(lab_t)) 

vis_cd4_crag <- sqldf(
  "select cd4d.*,crag.lab_dmy,crag.lab_t
    from vis_cd4_data cd4d
    left join lab_data_crag crag
      on cd4d.patient = crag.patient
        and crag.lab_dmy between(cd4d.cd4_date -7) and(cd4d.cd4_date +7)
  "
)%>%
  rename(
    crag_date = lab_dmy,
    crag_value= lab_t
  )

vis_cd4_crag<-vis_cd4_crag %>%
  arrange(desc(crag_date))

#Unique visits
#vis_cd4_crag_data_dup<-vis_cd4_crag[duplicated(vis_cd4_crag[,1:19]),]
vis_cd4_crag_data<-vis_cd4_crag[!duplicated(vis_cd4_crag[,1:19]),]


###################### TBLAM LAB DATA ##############################
####################################################################

lab_data_tblam <-lab %>%
  filter((lab_id =="LAM") & !is.na(lab_t)) 

vis_cd4_crag_tbm <- sqldf(
  "select cd4crag.*,tb.lab_dmy,tb.lab_t
    from vis_cd4_crag_data cd4crag
    left join lab_data_tblam tb
      on cd4crag.patient = tb.patient
        and tb.lab_dmy between(cd4crag.cd4_date -7) and(cd4crag.cd4_date +7)
  "
)%>%
  rename(
    tbm_date = lab_dmy,
    tbm_value= lab_t
  )

vis_cd4_crag_tbm<-vis_cd4_crag_tbm %>%
  arrange(desc(tbm_date))

#Unique visits
#vis_cd4_crag_tbm_data_dup<-vis_cd4_crag_tbm[duplicated(vis_cd4_crag_tbm[,1:21]),]
vis_cd4_crag_tbm_data<-vis_cd4_crag_tbm[!duplicated(vis_cd4_crag_tbm[,1:21]),]


###################### LAST VL DATA #################################
#####################################################################

lab_data_vl <-lab %>%
  filter((lab_id =="RNA") & !is.na(lab_v))

#Get last VL up to 15 months(390 days) before current visit
#Assumption: Patient under care should have VL atleast once a year
vis_cd4_crag_tbm_lvl<-sqldf(
  "select data.*,vl.lab_dmy,vl.lab_v
    from vis_cd4_crag_tbm_data data
    left join lab_data_vl vl
      on data.patient = vl.patient
        and vl.lab_dmy between (data.visit_dmy -390) and (data.visit_dmy -1)"
)%>%
  rename(
    lastvl_date = lab_dmy,
    lastvl_value= lab_v
  )

vis_cd4_crag_tbm_lvl<-vis_cd4_crag_tbm_lvl %>%
  arrange(desc(lastvl_date))

#Unique visits
#vis_cd4_crag_tbm_lvl_data_dup<-vis_cd4_crag_tbm_lvl[duplicated(vis_cd4_crag_tbm_lvl[,1:23]),]
vis_cd4_crag_tbm_lvl_data<-vis_cd4_crag_tbm_lvl[!duplicated(vis_cd4_crag_tbm_lvl[,1:23]),]


###################### LAST CD4 DATA #################################
######################################################################

#Get last CD4 for patient
#Assumption: CD4 within 3 months of current visit may still be valid

vis_cd4_crag_tbm_lvl_l3cd4<-sqldf(
  "
    select data.*,cd4.lab_dmy,cd4.lab_v
    from vis_cd4_crag_tbm_lvl_data data
    left join lab_data_cd4 cd4
      on data.patient = cd4.patient
        and cd4.lab_dmy between (data.visit_dmy -90) and (data.visit_dmy-1)
  "
)%>%
  rename(
    last3cd4_date = lab_dmy,
    last3cd4_value= lab_v
  )

vis_cd4_crag_tbm_lvl_l3cd4<-vis_cd4_crag_tbm_lvl_l3cd4 %>%
  arrange(desc(last3cd4_date))

#Unique visits
#vis_cd4_crag_tbm_lvl_l3cd4_data_dup<-vis_cd4_crag_tbm_lvl_l3cd4[duplicated(vis_cd4_crag_tbm_lvl_l3cd4[,1:25]),]
vis_cd4_crag_tbm_lvl_l3cd4_data<-vis_cd4_crag_tbm_lvl_l3cd4[!duplicated(vis_cd4_crag_tbm_lvl_l3cd4[,1:25]),]


################## LAST NEXT APPOINTMENT DATA ########################
######################################################################

vis_cd4_crag_tbm_lvl_l3cd4_lstnxtapp_data <- vis_cd4_crag_tbm_lvl_l3cd4_data %>%
  arrange(patient,visit_dmy)%>%
  group_by(patient) %>%
  mutate(# Get previous last next appointment and missed appointment days for patient
    last_next_app = lag(next_visit_dmy),
    miss_app_days = visit_dmy - last_next_app
  ) %>%
  select(last_next_app,visit_dmy,miss_app_days,everything())


########################## TRANSFER DATA #############################
######################################################################
#Add Transfers information to visit data
vis_cd4_crag_tbm_lvl_l3cd4_lstnxtapp_trans <-vis_cd4_crag_tbm_lvl_l3cd4_lstnxtapp_data%>%
  left_join(transfers,c("patient"="patient","visit_dmy"="transfer_dmy"))

vis_cd4_crag_tbm_lvl_l3cd4_lstnxtapp_trans<-vis_cd4_crag_tbm_lvl_l3cd4_lstnxtapp_trans %>%
  arrange(desc(visit_dmy))%>%
  #Move episode_id at the end
  relocate(episode_id,.after = last_col())

#Remove duplicate lines for transfers
#vis_cd4_crag_tbm_lvl_l3cd4_lstnxtapp_trans_data_dup<-vis_cd4_crag_tbm_lvl_l3cd4_lstnxtapp_trans[duplicated(vis_cd4_crag_tbm_lvl_l3cd4_lstnxtapp_trans[,1:31]),]
#Note: This will include transfers in and transfers out that happened the same day as two separate lines
vis_cd4_crag_tbm_lvl_l3cd4_lstnxtapp_trans_data<-vis_cd4_crag_tbm_lvl_l3cd4_lstnxtapp_trans[!duplicated(vis_cd4_crag_tbm_lvl_l3cd4_lstnxtapp_trans[,1:31]),]

#Adding a transfer category, transfer in or out
vis_cd4_crag_tbm_lvl_l3cd4_lstnxtapp_trans_data<- vis_cd4_crag_tbm_lvl_l3cd4_lstnxtapp_trans_data%>%
  mutate(
    trans_cat = case_when(
      mapply(identical,facility,facility_from)  ~ "transfer out",
      mapply(identical,facility,facility_to)  ~ "transfer in",
      TRUE ~ ""
    )
  )

```

##ELIGIBILITY CRITERIA
Note: It is important to note that some of the below eligibility criteria are not mutually exclusive, meaning a patient can belong to more than one eligibility criteria. As example, a patient can be returnee and has a high last VL. Important to clarify the order.  
-- Patients are considered and not visits events, meaning only first occurence of patient in the reporting period is considered
In the meantime, the following are assumptions for the eligibility:
New initiations : Patients who are starting ART 
  Exclusion : Transfers In
Returnees : Patients who are returning to care after being LTFUP (exact duration depends on settings/projects/countries)
High VL: Patients who visit the facility after having a high VL  
  Exclusion : Returnees

```{r INCLUDE=FALSE, fig.width=7.0}

######################################################################
####################### ELIBILITY CRITERIA ###########################
######################################################################



start_period <-as.Date("2021/01/01")
end_period <-as.Date("2021/12/31")
pat_col_ind <- 4 #Patient column index in the data frame
days_ltfup <-90 #This can be context specific

vis_cd4_crag_tbm_lvl_l3cd4_lstnxtapp_trans_data <- vis_cd4_crag_tbm_lvl_l3cd4_lstnxtapp_trans_data %>%
  mutate(
    vis_quarter_year = quarter(visit_dmy, with_year = T),
    vis_month_year   = format(as.Date(visit_dmy), "%Y-%m"),
  )

#Add category for the eligibility criteria
vis_cd4_crag_tbm_lvl_l3cd4_lstnxtapp_trans_data<-vis_cd4_crag_tbm_lvl_l3cd4_lstnxtapp_trans_data%>%
  mutate(
   elig_cat = case_when(
     visit_dmy == haart_dmy & method_into_art == 0 ~ "Initiations",
     
   ) 
  )


############################  PATIENTS WITH NO CD4   ###########################
################################################################################
#Q/ What time window between enrolment in cohort vs check for CD4?
#Joining patient table and lab table(CD4), then filter only patients with no CD4

#A/ Eligibility : Patients on ART for >=12M at start of reporting period and are still in care

no_cd4_pat <- sqldf(
  #Patient registered in Tier but have never had a CD4
  "
    select pat.*
    from pat
    where pat.patient not in (
      select patient from lab where lab_id ='CD4A'
    )
  "
) %>%
  filter(
    #Filter patients still active in care
    outcome=="20" &
      #Filter patients who started ART before end of reporting period and are on ART for more than 1 year
    (start_period - haart_dmy) >=365 #Difference report start >=12M
  )
#Join with visit table to get patient visits
no_cd4_pat<-no_cd4_pat %>%
  left_join(
    vis,c("patient" = "patient")
  ) %>%
  arrange(desc(visit_dmy))

#Keep only latest visit for each patient
no_cd4_pat_data<-no_cd4_pat[!duplicated(no_cd4_pat[,1:15]),]
no_cd4_pat_data %>%
  rename(#Renaming columns
    "lat_vis_dmy"      = visit_dmy,
    "lat_next_app_dmy" = next_visit_dmy
  )



###########################  NEW INITIATIONS   #################################
################################################################################



#Filter visits that happen on ART Start Date (new art initiations) and new patients (not transfers in)
new_art_initiations <-vis_cd4_crag_tbm_lvl_l3cd4_lstnxtapp_trans_data %>%
  filter( 
    visit_dmy == haart_dmy & method_into_art == 0
  )


art_casc <-vis_cd4_crag_tbm_lvl_l3cd4_lstnxtapp_trans_data %>%
  filter(visit_dmy>=start_period & visit_dmy<=end_period)%>%
   mutate(
    quarter_year = quarter(visit_dmy, with_year = T),
    month_year   = format(as.Date(visit_dmy), "%Y-%m"),
  )%>%
  
  group_by(quarter_year,facility)%>%
  summarise(
    tot_vis       = n(),
    tot_cd4       = sum(!is.na(cd4_value),na.rm = TRUE),
    tot_in        = sum(visit_dmy == haart_dmy & method_into_art == 0,na.rm = TRUE),
    cd4_in        = sum(visit_dmy == haart_dmy & method_into_art == 0 & !is.na(cd4_value),na.rm = TRUE),
    `% cd4_in`    = round((cd4_in/tot_in),4)*100,
    tot_cd4_lt200 = sum(cd4_value<200,na.rm = TRUE), #Tot CD4 < 200
    cd4_lt200_in  = sum(visit_dmy == haart_dmy & method_into_art == 0 & !is.na(cd4_value) & cd4_value<200,na.rm = TRUE), #Tot CD4 < 200
    cd4_lt200_oth = sum(!(visit_dmy == haart_dmy & method_into_art == 0) & !is.na(cd4_value) & cd4_value<200,na.rm = TRUE),
    tot_lam       = sum(cd4_value<200 & !is.na(tbm_value),na.rm = TRUE),
    `%tot_lam`    = round((tot_lam/tot_cd4_lt200),4)*100,
    lam_in        = sum(visit_dmy == haart_dmy & method_into_art == 0 & !is.na(cd4_value) & cd4_value<200 & !is.na(tbm_value),na.rm = TRUE),
    `%lam_in`     = round((lam_in/cd4_lt200_in),4)*100,
    tot_crag      = sum(cd4_value<200 & !is.na(crag_value),na.rm = TRUE),
    `%tot_crag`   = round((tot_crag/tot_cd4_lt200),4)*100,
    crag_in       = sum(visit_dmy == haart_dmy & method_into_art == 0 & !is.na(cd4_value) & cd4_value<200 & !is.na(crag_value),na.rm = TRUE),
    `%crag_in`    = round((crag_in/cd4_lt200_in),4)*100,
    lam_pos       = sum(cd4_value<200 & !is.na(tbm_value) & (tbm_value=="+"),na.rm = TRUE),
    `%lam_pos`    = round((lam_pos/tot_lam),4)*100,
    crag_pos      = sum(cd4_value<200 & !is.na(tbm_value) & (crag_value=="+"),na.rm = TRUE),
    `%crag_pos`   = round((crag_pos/tot_crag),4)*100,
  )# %>%
  #adorn_totals()


###################################  RETURNEES   ###############################
################################################################################
#Def: These are people who have been LTFUP and are coming back to care


returnees<- vis_cd4_crag_tbm_lvl_l3cd4_lstnxtapp_trans_data %>%
  filter(
    (visit_dmy>=start_period & visit_dmy<=end_period)
    & miss_app_days >days_ltfup)
  
#Filter to get unique patients in reporting period
#Get first occurrence of patient in reporting period
returnees <-returnees%>%
  arrange(patient,visit_dmy)

returnees_dup <-returnees[duplicated(returnees[,4]),]
returnees_data<-returnees[!duplicated(returnees[,4]),]

returnees_casc<-returnees_data%>%
  #Exclude initiations that might be part of the data (cross checking for data quality)
  filter( 
    !(visit_dmy == haart_dmy & method_into_art == 0)
  )%>%
  #Exclude patients who do not currently have a CD4 but have valid CD4 (<=3M)
  filter(
    !(is.na(cd4_value) & !is.na(last3cd4_value))
  )%>%
  mutate(
    quarter_year = quarter(visit_dmy, with_year = T),
    month_year   = format(as.Date(visit_dmy), "%Y-%m"),
  )%>%
  group_by(quarter_year)%>%
  summarise(
    tot_hvl        = n(),
    cd4_hvl        = sum(!is.na(cd4_value)),
    `% cd4_hvl`    = round((cd4_hvl/tot_hvl),4)*100,
  )


#############################  HIGH VIRAL LOAD   ###############################
################################################################################

#Visits with last High VL

#Q/ Consider people with valid CD4 as still part of the cascade for tblam and crag


high_vl <-vis_cd4_crag_tbm_lvl_l3cd4_lstnxtapp_trans_data %>%
  filter(lastvl_value >=1000) %>%
  arrange(patient,visit_dmy)

#Unique patients
high_vl_dup<-high_vl[duplicated(high_vl[,pat_col_ind]),]
high_vl_data<-high_vl[!duplicated(high_vl[,pat_col_ind]),]
#Get first occurrence of patient in reporting period
#Q/What happens if you had two high VL in same reporting period?


high_vl_casc<-high_vl_data%>%
  #Exclude initiations that might be part of the data (cross checking for data quality)
  filter( 
    !(visit_dmy == haart_dmy & method_into_art == 0)
  )%>%
  #Exclude patients who do not currently have a CD4 but have valid CD4 (<=3M)
  filter(
    !(is.na(cd4_value) & !is.na(last3cd4_value))
  )%>%
  mutate(
    quarter_year = quarter(visit_dmy, with_year = T),
    month_year   = format(as.Date(visit_dmy), "%Y-%m"),
  )%>%
  group_by(quarter_year)%>%
  summarise(
    tot_hvl        = n(),
    cd4_hvl        = sum(!is.na(cd4_value)),
    `% cd4_hvl`    = round((cd4_hvl/tot_hvl),4)*100,
  )


################### ELIGIBILITY FOR CD4 AND CD4 DONE ###########################
################################################################################

#Data for a year,show of trends
art_casc_sum1 <-art_casc%>%
  mutate(
    not_cd4_in = tot_in - cd4_in
  )%>%
  pivot_longer(cols=c(not_cd4_in,cd4_in))

art_casc_sum1 %>%
  mutate(
    name = factor(name,levels=c("not_cd4_in","cd4_in"))
  )%>%
  ggplot()+
  geom_bar(
    aes(x=month_year,y=value,fill=name),
    stat="identity",
    color = "#661a00"
  )+theme(
    #aspect.ratio = 1.3,
    legend.position = "bottom",
    axis.text.x = element_text(angle = 65,hjust = 1.0),
    legend.title = element_blank(),
  )+labs(
    title = "CD4 Done among new ART initiations, 2021",
    x = "Periode",
    y="Total"
  )+
  scale_fill_discrete(
    labels=c("","CD4 Done")
    )

#At a specific point in time
art_casc_sum2 <-art_casc%>%
 
  rename(
    "Eligible" = tot_in,
    "CD4 done" = cd4_in,
    "CD4 < 200"= cd4_lt200_in
  ) %>%
  pivot_longer(cols=c(`Eligible`,`CD4 done`,`CD4 < 200`))%>%
  mutate(
    name = factor(name,levels=c("Eligible","CD4 done","CD4 < 200"))
  )%>%
  mutate(
    category="ART Initiation"
  )

art_casc_sum2%>%
  ggplot()+
  geom_bar(
    aes(x=name,y=value,fill=category),
    stat="identity"
  )+theme(
    aspect.ratio = 0.8,
    #legend.position = "bottom",
    #axis.text.x = element_text(angle = 65,hjust = 1.0),
    legend.title = element_blank(),
  )+labs(
    title = "CD4 done by eligibility criteria, 2021",
    y="Total"
  )
  


```


```{e INCLUDE=FALSE}

########################### RETURNEES   ##############################
######################################################################
#Outcome : 10, Death(HIV Related), 11: Death (HIV relationship unknown), 20: Alive and in care, 30,31: Transfers out
#40,41: LTFUP, 95: Not ascertained

#Function to return last next appointment date for a patient
#based on patient id and current visit
last_next_app_fnc <- function(patient_id, curr_vis_date){
  
  vis_dmy  <- as.Date(curr_vis_date)
  vis_dmy_2<- seq(vis_dmy, length=2, by="-2 years")[2] #Two years from current visit

  
  result <- vis_cd4_crag_tbm_data %>%#Visit for the last 5 years
    filter(
      (patient == patient_id) &
      (visit_dmy < curr_vis_date &  visit_dmy >=vis_dmy_2) #Filter visits before current visit date up to 2 years ago
    ) %>%
    arrange(desc(visit_dmy)) %>% #Sort by date of visit
    select(next_visit_dmy) %>% 
    head(1) #Only get the first top visit

  result <- ifelse(nrow(result)==0,NA,result)
  value <- result[[1]]
  return (value)
}


eligible_vis <- vis_cd4_crag_tbm_data %>%
  filter(
    visit_dmy >=as.Date("2021/01/01") & visit_dmy <=as.Date("2021/12/31")
  )%>%
  left_join(transfers,
            c("patient" = "patient", "visit_dmy" = "transfer_dmy")) %>%
  filter(#Excludes all transfers (in and out)
    is.na (facility_from)) %>% 
  mutate(#Get last next appointment date
    last_next_app_date =  (map2(patient, visit_dmy, last_next_app_fnc)))
    

# eligible_vis_data <- sqldf(
#   "select el.* from eligible_vis el
#     left join (select max(vis.visit_dmy) as last_app_date
#                   from vis_cd4_crag_tbm_data vis
#                     where vis.patient = el.patient and vis.visit_dmy <el.visit_dmy)  
#   "
# )
# 
# mysql <-sqldf(
#   "select max(vis.visit_dmy) as last_app_date
#                   from vis_cd4_crag_tbm_data vis
#                     where vis.patient = '11ec5d34-78ed-41c9-b156-49d2140fda47' and vis.visit_dmy < '04/03/2020'
#   "
# ) 

vis_dmy  <- as.Date("2022/01/01")
  vis_dmy_2<- seq(vis_dmy, length=2, by="-14 month")[2] 
  vis_dmy_2
  
  

```

